---
title: "Data construction IMDB"
author: "Team 1"
date: "2025-02-05"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction
In today’s entertainment landscape, whether a TV series will be renewed for a second season is relevant and intriguing. Understanding the factors influencing renewal decisions can provide valuable insights for content creators, streaming platforms, and researchers studying audience engagement.

# Research objectives
This research seeks to investigate whether popular genres have a higher likelihood of renewal, if the language of the title plays a role in renewal decisions, and the significance of average ratings. It might seem intuitive that higher ratings lead to renewal ((What’s Behind a Show Renewal, n.d.).

# Methodology
To explore these factors, we will conduct a logistic regression analysis. The renewal status will be the dependent variable, while genre, language of the title, and average rating will serve as the independent variable. According to Lee and Wang (2003), logistic regression is a useful method for analyzing binary variables because it models and predicts the probability of a specific outcome. This method is useful as it can handle both continuous and categorial predictors, making it versatile for various types of data. 

#Expected contributions
The findings of this study can contribute to both academic and industry discussions. By identifying patterns in renewal decisions, this research can aid content creators and streaming services make informed choices about future productions. Additionally, the automated and reproducible workflow ensures that this research can be a valuable resource for other students and the broader scientific community.
Deployment
For deployment, the results will be communicated through a PDF report, ensuring accessibility and clarity for potential users. The structured format will effectively present conclusions, making it easy to interpret key findings.

#sources

Lee, E. T., & Wang, J. W. (2003). Statistical methods for survival data analysis. In Wiley series in probability and statistics. https://doi.org/10.1002/0471458546
What’s behind a show renewal. (n.d.). Parrot Analytics. https://www.parrotanalytics.com/insights/whats-behind-a-show-renewal/





## Loading the data

We will load the needed data files from the internet. It is all information that is not to do with the people that work on the movies/series, but rather the characteristics of the movies/series themselves

```{r warning=FALSE, message=FALSE}

# Define required packages
required_packages <- c("glue", "readr", "dplyr", "ggplot2", "knitr")

# Install missing packages
new_packages <- required_packages[!(required_packages %in% installed.packages()[, "Package"])]
if (length(new_packages)) install.packages(new_packages, type = "binary")

# Suppress startup messages while loading packages
suppressPackageStartupMessages({
  library(glue)
  library(readr)
  library(dplyr)
  library(ggplot2)
  library(knitr)
  library(tidyr)
  
})
```



```{r warning=FALSE, message=FALSE}
# Downloading IMDb datasets
options(timeout = 600)
download.file("https://datasets.imdbws.com/title.akas.tsv.gz", "title.akas.tsv.gz")
download.file("https://datasets.imdbws.com/title.basics.tsv.gz", "title.basics.tsv.gz")
download.file("https://datasets.imdbws.com/title.episode.tsv.gz", "title.episode.tsv.gz")
download.file("https://datasets.imdbws.com/title.ratings.tsv.gz", "title.ratings.tsv.gz")

# Read and process the data using readr functions and dplyr
akas_data <- read_tsv("title.akas.tsv.gz", col_names = TRUE, na = "\\N")

basics_data <- read_tsv("title.basics.tsv.gz", col_names = TRUE, na = "\\N")

episode_data <- read_tsv("title.episode.tsv.gz", col_names = TRUE, na = "\\N")

ratings_data <- read_tsv("title.ratings.tsv.gz", col_names = TRUE, na = "\\N") 

```


## New dataset construction

We will first create a new dataset, as creating the variables per existing dataset is not preferred  because of the sizes of them.

The new dataset is called `TV_series_data`, and contains data on different TV_series. This dataset is constructed by matching the parenTconst ID's in the `episode_data` dataset to the other unique ID's in other the other datasets.


```{r warning=FALSE, message=FALSE}


# Merge episode_data with title.basics, title.akas, and title.ratings
TV_series_data <- episode_data %>%
  left_join(basics_data, by = c("parentTconst" = "tconst")) %>%
  left_join(akas_data, by = c("parentTconst" = "titleId")) %>%
  left_join(ratings_data, by = c("parentTconst" = "tconst"))

# Check the result
glimpse(TV_series_data)

```


```{r warning=FALSE, message=FALSE}
# check if titleType is only tvSeries
table(TV_series_data$titleType)

# see that there is also tvMiniSeries, we don't want to include them as TV Mini Series are usually meant to only last a season, and don't get renewed.


# Only keep TV_series_data
TV_series_data <- TV_series_data %>%
  filter(titleType == "tvSeries")

```

```{r}
# check unique nr of parentTconst in episode_data and TV_series_data

print(length(unique(episode_data$parentTconst)))   # 215686
print(length(unique(TV_series_data$parentTconst))) # 179223
```

```{r}
# only keep one row per parentTconst, and only keep the row with the highest seasonNumber of that parentTconst


# Select the row with the highest seasonNumber for each parentTconst
TV_series_data <- TV_series_data %>%
  group_by(parentTconst) %>%
  slice_max(seasonNumber, with_ties = FALSE) %>%
  ungroup()


head(TV_series_data)

print(nrow(TV_series_data)) # should be 179223

```


We will create the variables necessary for our analysis:
(1) Renewed = wheter a TV series was renewed for a second season
(2) Genre = the genres (encoded with label encoding)

Furthermore, we will use all other variables not used for the creation of these as variables in our logistic regression analysis.

These are combined into a single dataset, and the variables that are not included to construct the variables above, are also added and act as control variables in this research.


### Create Renewed

Add binary variable `Renewed` to the `TV_series_data` dataset:


- TV series with their highest `seasonNumber` >= 2 get 1
- TV series with their highest `seasonNumber` == 1 get 0


```{r , echo=FALSE}


TV_series_data <- TV_series_data %>%
  # Create the Renewed variable based on the season number being above or equal to 2
  mutate(Renewed = ifelse(!is.na(seasonNumber) & seasonNumber >= 2, 1, 0)) %>%
  # Ungroup to return to a standard data frame
  ungroup()

# Check the result
head(TV_series_data)
table(TV_series_data$Renewed)

```

### Create Genre

Genre = the genres (encoded with label encoding)

As the genres are in a comma-separated format, we will split them into separate columns.

```{r warning=FALSE, message=FALSE}


# Sample data

# Split the multi_values column into three separate columns
TV_series_data_genre <- TV_series_data %>%
  separate(genres, into = c("Genre1", "Genre2", "Genre3"), sep = ",", fill = "right")


```
Then, as we want to use them in analysis, we will encode them using label encoding.

```{r}

# Get all unique categories across the selected columns
all_genres <- unique(unlist(TV_series_data_genre))

# Apply the same factor levels consistently across columns
TV_series_data_genre <- TV_series_data_genre %>%
  mutate(across(starts_with("Genre"), 
                ~ as.numeric(factor(., levels = all_genres)), 
                .names = "{.col}_encoded"))



```


```{r warning=FALSE, message=FALSE}
# view the mapping of each variable

# Print out the mapping for each column
factor_mapping <- sapply(TV_series_data_genre[, c("Genre1", "Genre1_encoded")], function(x) {
  levels(factor(x))
})

print(factor_mapping)

```



We unfortunately found that all the values for the Language column are NA (missing), meaning that we cannot use this variable to construct the wanted binary variable that indicates whether the original title is in English or not.


```{r warning=FALSE, message=FALSE}
# to test wheter this is not an error, we check the original dataset where this variable was in to see if there are any values in the language column
# as can be seen below, this is the case
table(TV_series_data_genre$language)
```

# Data cleaning and EDA


## Variables that are available
In our dataset we have several variables available to us. In order to conduct a proper analysis we must start with defining what all of the variables mean and how they are calculated. We will derive the definition of these variables from the IMDB website from which we downloaded these sources:



### Existing variables in the dataset

**(1) tconst** - alphanumeric unique identifier of the title
**(2) parentTconst (string)** - alphanumeric identifier of the parent TV Series
**(3) seasonNumber (integer)** – season number the episode belongs to. In the case of our dataset the number of seasons is stored here (so 8 translates to 8 seasons).
**(4) episodeNumber (integer)** – episode number of the tconst in the TV series. In the case of our dataset episode number refers to the most recent episode released (so a if the most recent season is episode so if this was season 5 episode 3 for a given show, this would be denoted as 3)
**(5) titleType (string)** – the type/format of the title (e.g. movie, short, tvseries, tvepisode, video, etc)
**(6) titleprimaryTitle (string)** – the more popular title / the title used by the filmmakers on promotional materials at the point of release
**(7) originalTitle (string)** - original title, in the original language
**(8) isAdult (boolean)** - 0: non-adult title; 1: adult title
**(9) startYear (YYYY)** – represents the release year of a title. In the case of TV Series, it is the series start year
**(10) endYear (YYYY)** – TV Series end year. '\N' for all other title types
**(11) runtimeMinutes** – primary runtime of the title, in minutes
**(12) ordering (integer)** – a number to uniquely identify rows for a given titleId
**(13) title (string)** – the localized title

**(14) region (string)** - the region for this version of the title
**(15) language (string)** - the language of the title
types (array) - Enumerated set of attributes for this alternative title. One or more of the following: "alternative", "dvd", "festival", "tv", "video", "working", "original", "imdbDisplay". New values may be added in the future without warning
**(16) types (array)** - Enumerated set of attributes for this alternative title. One or more of the following: "alternative", "dvd", "festival", "tv", "video", "working", "original", "imdbDisplay". New values may be added in the future without warning
**(17) attributes (array**) - Additional terms to describe this alternative title, not enumerated
**(18)isOriginalTitle (boolean)** – 0: not original title; 1: original title
**(19) averageRating** – weighted average of all the individual user ratings
**(20) numVotes** - number of votes the title has received






### Our own generated variables
Lastly, in the data preparation process, we created our own varibales that we used for subsequent analysis. Below the variable name and what it entails is documented. 

**(21) renewed** - this variable encodes whether a given TV series has been renewed for a second season. This was deduced based on whether a season two was on file. When a TV show had more than one season, this was encoded as 1, when a TV show only had one season in the dataset, this was encoded as a 0. 

**Genre1, Genre2 and Genre3** - a problem with the raw data set was that the first, second and third genres were all collected in one big variable. This makes analysis of these genres quite challenging as this creates a very large number of unique combinations. Thus, for ease of analysis, it was decided to break up the string of genres into three different columns: genre 1, genre 2, genre 3. Genre 1 thus denotes the first genre tagged in the database, genre 2 the second and genre 3 the third.Though it seems logical to assume that the first genre tagged is the primary (or main) genre, we cannot operate under this assumption as this is not noted anywhere. 

**Genre1_encoded, Genre2_encoded and Genre3_encoded**
These variables were used to create Genre1, Genre 2 and Genre3 but by themselves do not mean much.

## Removing useless variables and taking a quick glimpse
A lot of variables in the dataset TV_series_data_genre is not information we need to keep for later analysis. Thus we start by removing the following variables, since they are not necessary for our analysis or are repetative:

**(1) SeasonNumber**: we are looking at whether a TV show is renewed in the first season or not, whether a TV show runs for 2 or 5 seasons is outside the scope of out study
**(2) Titletype**: all our title types are TV-series since this is what we want to be looking at exclusively
**(3) EpisodeNumber**: this variable does not tell as much, only the most recent episode number
**(4) RuntimeMinutes**: a tv show renewed for a second season will have a longer runtime, moreover a TV show that has been renewed several seasons will have a longer runtime. This is more a product of renewal, not something that influences renewal itself.
**(5) genre1_encoded, genre2_encoded, genre3_encoded.** : This is a numeric value used in the data preperation process but is not much use to us in the actual analysis
**(6) ordering**: used to identify rowsin one dataset but is not something that carries over in other datasets
**(7) tconst**: repetative information (was used in merging)

Okay so lets just review all the data we have by breaking down and looking at the summary and glimpse per review. We have already removes some columns that are no longer relevant. At this point in time we are keeping variables such as whether the tv show is targeted towards adults or not since we might want to use them later for a predictive analysis (possiby).

```{r}
summary(TV_series_data_genre)
TV_series_data_clean = TV_series_data_genre %>%
  select(c(-seasonNumber, -titleType, -attributes, -runtimeMinutes, -episodeNumber, -Genre1_encoded, -Genre2_encoded, -Genre3_encoded, -ordering, -tconst))

glimpse(TV_series_data_clean)
summary(TV_series_data_clean)
```


## A clear overview: 
Though the glimpse function is useful, lets create a custom overview, so that we can analyse the data in more detail. We will create 3 table overviews:

**1) data_types_summary**: This table will include three different columns that summarises the following about TV_series_data_clean
1.1 Variable: Name of the variable
1.2 type: type of variable that is stored (numeric, string etc.)
1.3 Percentage_na: the percentage of NA's for this variable in our dataset

**2) data_types_numeric**
2.1 Variable: Name of the variable
2.2 minimum: the lowest value for a given variable
2.3 maximum: the highest value for a given variable
2.4 average_mean: the mean average  for a given variable
2.5 average_median: the median average  for a given variable
2.6 average_mode: the mode average  for a given variable
2.7 binary: checks if the only unique values are 1 or 0. if this is true it is likely this variable is one for which a condition is either true or false


**3) data_types_char**
3.1 Variable: Name of the variable 
3.2 unique_values: checks the number of unique character values, might be useful with plotting (for example if a given variable has over 1000 unique variables versus 8, this might influence how we choose to plot or visually represent them)


```{r}
## find all the columns we  have in the first place
columns = colnames(TV_series_data_clean)



## create 3 empty dataframes. the first one stores whether a vector is a character or numeric. then based on this 
data_types_summary = data.frame(variable = character(), type = character(), percentage_na = character(), stringsAsFactors = FALSE)
data_types_numeric = data.frame(variable = character(), minimum = numeric(), maximum = numeric(), average_mean = numeric(), average_median = numeric(), average_mode = numeric(), binary = numeric(), stringsAsFactors = FALSE)
data_types_char = data.frame(variable = character(), unique_values = character(), stringsAsFactors = FALSE)

## We could not find a mode average function in r so we got this one from online 
## https://www.tutorialspoint.com/r/r_mean_median_mode.htm 

getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

## Create a function whether a variable is binary 

binary_check = function(v) {
  if (all(v %in% c(0, 1))) {
  return(1)
  } else { 
  return(0)
    }
  
}




for (col in columns) {
  
  column_type = class(TV_series_data_clean[[col]])
  amount_na = (mean(is.na(TV_series_data_clean[[col]]))) * 100
  
  ## df1: The type of data from our main dataframe
  data_types_summary <- rbind(data_types_summary, data.frame(variable = col, type = column_type, percentage_na = amount_na, stringsAsFactors = FALSE))
  
  if (column_type == "numeric") {
    
    ## DF2: if a value is numeric the min and max value will be collected for this variable 
    min_value = min(TV_series_data_clean[[col]], na.rm = T)
    max_value = max(TV_series_data_clean[[col]], na.rm = T)
    mean_value =  mean(TV_series_data_clean[[col]], na.rm = T)
    median_value = median(TV_series_data_clean[[col]], na.rm = T)
    mode_value = getmode(TV_series_data_clean[[col]])
    binary_value = binary_check(TV_series_data_clean[[col]])

    
    data_types_numeric <- rbind(data_types_numeric, data.frame(variable = col, minumum = min_value, maximum = max_value, average_mean = mean_value, average_median = median_value, average_mode = mode_value, binary = binary_value, stringsAsFactors = FALSE))
    
  } else {
    
    ## DF3: if a value is not numeric it has to be character so the number of NAs are collected 
    number_unique_var = length(unique(TV_series_data_clean[[col]]))

    data_types_char <- rbind(data_types_char, data.frame(variable = col, unique_values = number_unique_var, stringsAsFactors = FALSE))
    
  }
  
  
}


kable(data_types_summary, caption = "Data Types Summary")
kable(data_types_numeric, caption = "Data Types Summary: Numeric Only")
kable(data_types_char, caption = "Data Types Summary: only Characters")
```

###### New Code from (329-399) below
```{r}
## Get all column names
columns <- colnames(TV_series_data_clean)

## Initialize empty lists instead of inefficient data frames
data_types_summary_list <- list()
data_types_numeric_list <- list()
data_types_char_list <- list()

## Mode function (Ensures it works correctly)
getmode <- function(v) {
  uniqv <- unique(na.omit(v)) # Remove NAs before computing mode
  if (length(uniqv) == 0) return(NA) # Handle empty case
  uniqv[which.max(tabulate(match(v, uniqv)))]
}

## Binary check function
binary_check <- function(v) {
  all(v %in% c(0, 1)) * 1 # Returns 1 if binary, else 0
}

## Loop through each column
for (col in columns) {
  column_data <- TV_series_data_clean[[col]]
  column_type <- class(column_data)
  amount_na <- mean(is.na(column_data)) * 100

  ## Store data type summary
  data_types_summary_list[[col]] <- data.frame(
    variable = col, type = column_type, percentage_na = amount_na, stringsAsFactors = FALSE
  )

  if (is.numeric(column_data)) {
    ## Compute numeric statistics
    min_value <- min(column_data, na.rm = TRUE)
    max_value <- max(column_data, na.rm = TRUE)
    mean_value <- mean(column_data, na.rm = TRUE)
    median_value <- median(column_data, na.rm = TRUE)
    mode_value <- getmode(column_data)
    binary_value <- binary_check(column_data)

    ## Store numeric data
    data_types_numeric_list[[col]] <- data.frame(
      variable = col, minumum = min_value, maximum = max_value, 
      average_mean = mean_value, average_median = median_value, 
      average_mode = mode_value, binary = binary_value, stringsAsFactors = FALSE
    )
  } else {
    ## Store character data summary
    number_unique_var <- length(unique(na.omit(column_data))) # Ignore NAs
    data_types_char_list[[col]] <- data.frame(
      variable = col, unique_values = number_unique_var, stringsAsFactors = FALSE
    )
  }
}

## Convert lists to data frames
data_types_summary <- do.call(rbind, data_types_summary_list)
data_types_numeric <- do.call(rbind, data_types_numeric_list)
data_types_char <- do.call(rbind, data_types_char_list)

## Display results using kable
kable(data_types_summary, caption = "Data Types Summary")
kable(data_types_numeric, caption = "Data Types Summary: Numeric Only")
kable(data_types_char, caption = "Data Types Summary: Only Characters")
```


## variable inspection and subsequent course correction

Most of the ranges in the numeric table seems to be normal, however the variable startyear has values p to 2026. This makes sense as a production might have already added a TV series to IMDB in order to add trailers, even though the show has not started airing. However, this creates an unfair bias where more recent shows will be less likely to have been renewed, through virtue of not having finished airing or having had the opportunity to renew a show. In an analysis by parrot analytics they used a range of 80 days to determine whether a show would be renewed or not (Parrot Analytics, 2023). Since this data has been updated as of March 2024, its more fair to only include data up to 2023 (to avoid recency bias)

Moreover, isOriginalTitle is a variable for which all the values are 1, in other words all titles in the database are original. Thus for analysis purposes, it does not make sense to include this varibale in future analysis. The same goes for Types, for which the variable is either original or NA, which does not much either. 

Notwithstanding the Startyear and the NA's, the data seems to be in regular ranges, The variable types are also correct.


```{r}

TV_series_data_clean = TV_series_data_clean %>%
  filter(startYear < 2024)

```

###### New code (479-484) below
```{r}
TV_series_data_clean <- filter(TV_series_data_clean, startYear < 2024)

```


```{r}

TV_series_data_clean = TV_series_data_clean %>%
  select(-isOriginalTitle, -types)

```

###### New code (493-498) below
```{r}
TV_series_data_clean <- TV_series_data_clean[, !(names(TV_series_data_clean) %in% c("isOriginalTitle", "types"))]

```

## Looking at all the available NA variables

Okay so as mentioned earlier there are some NA's in the Dataset, for ones which the amount of na's is 100 percent will naturally have to be removed (as this means there is no data on them). as for the variables that do have an NAs it could be interesting to explore whether the NA data has different properties than the data which does not have NAs (for example if a lot of NA data on number of votes also has no renewals, this could influence how we interpret our final results)

First let show a bar plot of all the percentage of missing values.
```{r}

ggplot(data_types_summary, aes(x = reorder(variable, -percentage_na), y = percentage_na)) +
  geom_bar(stat = "identity") + 
  labs(title = "NA vs Non-NA Values for All Columns",
       x = "Type variable",
       y = "Percentage of NA values") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```

As was already evident in data_types_summary, and is visualized in the plot above, some variables have values for which all the information is missing, and thus can be removed. 

```{r}

## collecting variables for which there is no data (100 percent NA)
all_missing_data <- data_types_summary$variable[which(data_types_summary$percentage_na == 100)]

# Print the result
glue("missing all the data for: {all_missing_data}")




TV_series_data_clean <- TV_series_data_clean %>%
  select(-all_of(all_missing_data))


```

###### New code (525-540) below
```{r}
## Collect variables with 100% missing data
all_missing_data <- data_types_summary$variable[data_types_summary$percentage_na == 100]

# Print missing variables
if (length(all_missing_data) > 0) {
  glue("Missing all data for: {paste(all_missing_data, collapse = ', ')}")
} else {
  glue("No variables are completely missing data.")
}

## Remove columns with 100% missing data
TV_series_data_clean <- TV_series_data_clean[, !names(TV_series_data_clean) %in% all_missing_data]

```

## Looking at our NA distribution before removing it
HOw our NA's are distributed among the independent variable is of relevance to our analysis. After all, if we find Na's to be much more present in either conditions, the analysis performed might be limited by data storage or collection practices done by IMDB.

Though we can manually code each individual plot and their NA ratios per variable, looping through our variables can make our code less repetitive and makes the result less prone to human errors such as typo's.  


## All NA plots

```{r }

### values which in the summary had more than one percent NA's
target_columns = data_types_summary$variable[which(data_types_summary$percentage_na > 1)]

## TAKE out already removed columns

columns = colnames(TV_series_data_clean)

target_columns <- target_columns[target_columns %in% columns]

## As we have already filtered startyear (and discarded NA variables in this category) we will not be analysing it for this part

target_columns <- target_columns[target_columns != "startYear"]

target_columns <- target_columns[target_columns != "title"]



for (col in target_columns) {
  print(
    ggplot(TV_series_data_clean, aes(x = as.factor(Renewed), fill = as.factor(is.na(.data[[col]])))) +
      geom_bar(position = "dodge") +  
      labs(title = paste("Comparison of NA vs. Non-NA in", col, "and TV Show Renewal"),
           x = "Renewal Status",
           y = "Count",
           fill = "Category") +
      scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "blue"), 
                        labels = c(paste(col, "Data Available"), paste(col, "Data Missing"))) +
      theme_minimal()
  )
}

```

###### New code (567-600) below
```{r}
# Identify target columns with >1% NA values
target_columns <- data_types_summary %>%
  filter(percentage_na > 1) %>%
  pull(variable) %>%
  intersect(names(TV_series_data_clean)) %>%
  setdiff(c("startYear", "title"))  # Exclude specific columns

# Generate plots for each target column
plots <- lapply(target_columns, function(col) {
  ggplot(TV_series_data_clean, aes(x = factor(Renewed), fill = factor(is.na(.data[[col]])))) +
    geom_bar(position = "dodge") +
    labs(
      title = paste("Comparison of NA vs. Non-NA in", col, "and TV Show Renewal"),
      x = "Renewal Status",
      y = "Count",
      fill = "Category"
    ) +
    scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "blue"),
                      labels = c(paste(col, "Data Available"), paste(col, "Data Missing"))) +
    theme_minimal()
})

# Print all plots
print(plots)
```


## A table Summary With all the NAs
Though the plots are useful, it is also useful to have the exact percentages of the NAs for both renewed conditions to just see how much data is missing relatively (and not in terms of count). This has been added below using the table as follows:
(1) Variable: Name of variable
(2) na_perc_total: total percentage of NA's for a variable
(3) na_perc_ren: percentage of NA's for which a variable is renewed (1)
(4) na_per_ren:percentage of NA's for which a variable is not renewed (0)

```{r}


na_summary_df <- data.frame(
  Variable = character(),
  na_perc_total = numeric(),
  na_perc_ren = numeric(),
  na_perc_not_ren = numeric(),
  stringsAsFactors = FALSE
)

for (col in target_columns) {
  
  Na_count <- sum(is.na(TV_series_data_clean[[col]]))
  na_count_renewed <- sum(is.na(TV_series_data_clean[[col]]) & TV_series_data_clean$Renewed == 1)
  na_count_not_renewed <- sum(is.na(TV_series_data_clean[[col]]) & TV_series_data_clean$Renewed == 0)
  
  Na_percentage <- (Na_count / nrow(TV_series_data_clean)) * 100
  Na_percentage_renew <- (na_count_renewed / sum(TV_series_data_clean$Renewed == 1)) * 100
  Na_percentage_not_renew <- (na_count_not_renewed / sum(TV_series_data_clean$Renewed == 0)) * 100
  
  na_summary_df <- rbind(na_summary_df, data.frame(
    Variable = col,
    na_perc_total = Na_percentage,
    na_perc_ren = Na_percentage_renew,
    na_perc_not_ren = Na_percentage_not_renew
  ))
}

# Print final dataframe
kable(na_summary_df, caption = "Summary of percentage of missing variables in Total and Renewal Conditions")


```

###### New code (638-671) below
```{r}
# Compute NA percentages for total, renewed, and not renewed
na_summary_df <- target_columns %>%
  map_dfr(~ {
    col <- .x
    total_na <- mean(is.na(TV_series_data_clean[[col]])) * 100
    na_renewed <- mean(is.na(TV_series_data_clean[[col]])[TV_series_data_clean$Renewed == 1]) * 100
    na_not_renewed <- mean(is.na(TV_series_data_clean[[col]])[TV_series_data_clean$Renewed == 0]) * 100
    
    tibble(Variable = col, na_perc_total = total_na, na_perc_ren = na_renewed, na_perc_not_ren = na_not_renewed)
  })

# Display the result
kable(na_summary_df, caption = "Summary of percentage of missing values in Total and Renewal Conditions")

```


## Interpretation of the NA values

(1) Endyear: more of the Endyear data is missing for shows that have not been renewed. This could be because no Endyear has yet been determined (and thus not listed). This might be interesting to consider in the final analysis with our results.

(2) Genre 1: we can see that the percentage of shows which have no genre tag is greater in those who have not been renewed, notwithstanding the difference is not that large (3.96% missing in renewed shows, 7.32% in those which have not been renewed). 

(3) Genre2 & Genre3: though a missing Genre2 and Genre3 is not a indicative of anything (it could mean a show only has 1 tag) it is interesting to analyse. For example both for Genre2 and Genre3 have more missing values for shows which were not renewed (in other words renewed shows might have more tags). This could be interesting to look at in further analysis

(4) AverageRating: There is a lot more data missing for AverageRating in shows that have not been renewed, than for those who have. according to IMDB's own policy, it does not show ratings for shows which have less than 5 ratings. Perhaps the higher number average ratings are missing for shows that are not renewed is thus a byproduct of them not being popular. Ergo, are future analysis should note that shows that have less than 5 ratings were removed in the analysis process. 

(5) NumVotes: Again, this ties in with the average ratings. Perhaps a number of votes is also not noted when this number is too low.Again, by filter out the datapoints for which NA values are missing, we are also filtering out shows which are very unpopular on IMDB.  





## Removing the NA 
as our research is primarily concerned with renewals,average rating, popularity (lanauge is no longer in the dataset so that cant be studied), lets remove the data for which we have NA's for average rating and popularity (num votes)

```{r}


TV_series_data_clean = TV_series_data_clean %>%
  filter(!is.na(averageRating) & !is.na(numVotes))

head(TV_series_data_clean)
```




## Plotting our variables (numeric)
Now that we have a Dataset where all the NA values are removed, lets plot the different variables. For our numeric variables we are going to be using Barplots and Boxplots. For variables that only have a value of  0 or 1 (binary or boolean), it does not make sense to create a Boxplot since the distribution is not continuous or ordinal. For variables that are not binary boxplots make more sense as we can glean quartiles, medians and extreme values quite easily.

As we already have information on whether a variable is binary from our previous analysis, we can use this information again to create our plots without manually coding the boxplot or the barplot depending on what type of variable we are dealing with

```{r}

## these are all our numeric variables
remaining_columns_sans_genre = c("isAdult", "startYear", "endYear", "averageRating", "numVotes", "Renewed")

for (col in remaining_columns_sans_genre) {
  ## IF the factor was previously identified as numeric
  
  data_type = data_types_summary[which(data_types_summary$variable == col),"type"]

  
  if (data_type == "numeric") {
    
    binary = data_types_numeric[which(data_types_numeric$variable == col),"binary"]
    
    if (binary == 0) {
      
      boxplot(TV_series_data_clean[[col]], main = paste("Boxplot of", col), ylab = col)
      
      
      
      
      
    } else {
      
      
      barplot(table(TV_series_data_clean[[col]]), main = paste("Barplot of", col), xlab = col, ylab = "Count", col = c("red", "green"))
      
    }
      
      
    
    
    
  }
    

}



```

###### New code (728-769) below (#1) -> Idk which one is best, see the following code
```{r}
# Define numeric variables of interest
remaining_columns_sans_genre <- c("isAdult", "startYear", "endYear", "averageRating", "numVotes", "Renewed")

# Filter numeric variables from the summary table
numeric_vars <- data_types_summary %>%
  filter(variable %in% remaining_columns_sans_genre, type == "numeric") %>%
  pull(variable)

# Generate plots based on binary classification
plots <- lapply(numeric_vars, function(col) {
  is_binary <- data_types_numeric %>%
    filter(variable == col) %>%
    pull(binary) %>%
    as.logical()
  
  if (is_binary) {
    barplot(table(TV_series_data_clean[[col]]),
            main = paste("Barplot of", col),
            xlab = col, ylab = "Count",
            col = c("red", "green"))
  } else {
    boxplot(TV_series_data_clean[[col]], main = paste("Boxplot of", col), ylab = col)
  }
})

# Print all plots
print(plots)
```
###### New code (728-769) below (#2) -> Idk which one is best
```{r}
# Define numeric variables of interest
remaining_columns_sans_genre <- c("isAdult", "startYear", "endYear", "averageRating", "numVotes", "Renewed")

# Filter numeric variables from the summary table
numeric_vars <- data_types_summary %>%
  filter(variable %in% remaining_columns_sans_genre, type == "numeric") %>%
  pull(variable)

# Generate and print plots using ggplot2
for (col in numeric_vars) {
  is_binary <- data_types_numeric %>%
    filter(variable == col) %>%
    pull(binary) %>%
    as.logical()
  
  plot <- if (is_binary) {
    ggplot(TV_series_data_clean, aes(x = as.factor(.data[[col]]))) +
      geom_bar(fill = c("red", "green")) +
      labs(title = paste("Barplot of", col), x = col, y = "Count") +
      theme_minimal()
  } else {
    ggplot(TV_series_data_clean, aes(y = .data[[col]])) +
      geom_boxplot(fill = "blue", alpha = 0.5) +
      labs(title = paste("Boxplot of", col), y = col) +
      theme_minimal()
  }
  
  print(plot)  # Ensure each plot is displayed properly
}

```


### interpretation of plots:
(1) barplot isAdult: Is is clearly evident that the number of TV-series that have an adult rating is significantly smaller than those who do not have an adult rating
(2) Startyear: there are significant more TV-series that have started airing in the early to late 2000s than the late 1900 (Negatively skewed). This makes sense as IMDB has only been around since 1990, and showrunners from that time might be less concerned about adding their tv Show. Moreover, given IMDBs recent popularity, this might also be considered a birgger priority for current shows. 
(3) Endyear: we observe the same thing for Endyear as we did for Startyear, this makes sense for the same reasoning
(4) Averagerating: averageRating looks normal but is slightly negatively skewed, indicating that people are in general a positive (as otherwise 5 would be the average)
(5) Numvotes:NumVotes is quite positively skewed, indicating that most shows receive a (relative) low number of votes and a few shows get a very large number of votes. In other words, popularity is not discributed normally. 


## GENRE encoding 
now lets look at genre. As a TV show can have several Genre tags we cannot simply plot a single variable but use pivot longer to count the percentage of occurences of a genre in total. Thus in the table below percentage indicates the percentage of times a genre is tagged in general (not the percentage of times a genre is tagged for a TV show).

```{r}


genre_counts <- TV_series_data_clean %>%
  select(Genre1, Genre2, Genre3) %>%  
  pivot_longer(cols = everything(), names_to = "Genre_Type", values_to = "Genre") %>%  
  drop_na() %>% 
  count(Genre, sort = TRUE)  

# Display the result

genre_counts = genre_counts %>%
  mutate(percentage = (n/sum(n)*100))

kable(genre_counts, caption = "Frequency of Genres occurences in our dataset")
```

###### New code (847-862) below
```{r}
# Count occurrences of each genre
genre_counts <- TV_series_data_clean %>%
  select(Genre1, Genre2, Genre3) %>%
  pivot_longer(everything(), names_to = "Genre_Type", values_to = "Genre") %>%
  drop_na() %>%
  count(Genre, sort = TRUE) %>%
  mutate(percentage = (n / sum(n)) * 100)

# Display the result
kable(genre_counts, caption = "Frequency of Genre Occurrences in Dataset")

```


as is evident from the table and the graph below, the most common genre is comedy, and drama, whereas western, war and musical are less popular
```{r}

ggplot(genre_counts, aes(x = reorder(Genre, -n), y = n, fill = Genre)) +
  geom_bar(stat = "identity") +
  labs(title = "TV Genre Distribution", x = "Genres", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(fill = "none") 

```

###### New code (881-890) below
```{r}
ggplot(genre_counts, aes(x = reorder(Genre, -n), y = n, fill = Genre)) +
  geom_col() +
  labs(title = "TV Genre Distribution", x = "Genres", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  guides(fill = "none")
```

